---
title: "WUP_04"
author: "Marina Ornelas"
date: "Spring 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
mydata <- read.csv(file = "GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey. The question to answer is to resolve this hypothesis.\
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE\
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(mydata,aes(x=DEGREE,y=WORDSUM))

basicCN + geom_boxplot()
```

This box shows that both High school and and Lt High school participants would likely end up below the results of participants in the Graduate category.We have learned that when we have boxes apart from each other if they were to be placed in the same column this is generally an indicator that the data is different.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=mydata)
```

The data shows that the average for the Graduate degree is 10.10 being the highest, then Junior College being 9.92 as the second highest, then Bachelor Degree being 9.20 being the third, the fourth being High School with the average of 8.28 and 
Lt High School being the lowest with an average of 7.75. The Q1 value for graduates is a 10 while the high school and lt high school are values for Q3 are 9 and 8.25 making them lower. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=mydata)
summary.aov(model1)
```

The probability of this data not having a difference is 0.000236%. This tells us that there is a difference but we have to run multiple comparison tests to find out the differences and figure out exactly what categories differ with each other. 

```{r}
a <- TukeyHSD(model1, conf.level=.99)
par(mar=c(3,15.7,2,2))
plot(a,las=2)
```




# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
